---
title: "Final Project Proposal"
author: "Laura Jiménez"
date: "April, 2018"
output: html_document
fontsize: 11pt
---

<!--- Chunk for calling any extra code needed in this document such as libraries, functions or data --->
```{r extra-code,include=FALSE}
# used in chunk "data"
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) {X = file.info(x)$mtime})
}
source("plotdata.R")
```

## The problem

  My current research focuses on the problem of estimating the hypothetical fundamental niche of a species from presence data. The method that I'm developing estimates a niche by:  
1. Defining an environmental space in which each axis represents a climatic variable. In this space we can plot the existing environmental conditions from the geographic region of interest and use them mainly for interpretating the results.  
2. Using georeferenced occurrences of the species and its corresponding environmental conditions as the primary data, and climatic tolerance ranges from physiological experiments as _a priori_ information.  
3. Postulating a simple shape for the niche of a species (Hutchinson's fundamental niche). I use a multivariate normal model to represent the fundamental niche, so that the parameters of interest are a vector of means, $\mu$, and a precision matrix, $A$ (which is defined as the inverse of the covariance matrix).  
4. Providing a probabilistic (i.e. Bayesian) statement of the problem and the posterior distribution for the parameters that define the fundamental niche.  

## Purpose and end products of the project

  The ultimate goal of the final project will be to create an R package to perform all the calculations needed to apply this method of niche estimation and to analyze the results.  
  The R package will contain all the functions that the users will need to apply the model to their data, to produce visualizations of their data and the results of the niche estimation, and to adapt the model to different variations in the data.  
  The second product of this final project will be a manual that contains a description of the functions and datasets included in the package, together with some examples that explain how to use them.

## A Bayesian method for niche estimation

  Three kinds of data are needed to apply our method:  
1. __Background data__: measurements of the environmental variables that are assumed to be the most important for describing the fundamental niche of the species of interest. This set of environmental conditions correspond to all the geographical sites belonging to a grid that covers a geographic region of reference.  
2. __Occurrence data__: georeferenced presences of the species and its corresponding measurements of environmental variables.  
3. __Tolerance ranges__: intervals of tolerance for each variable that define the environemental variable, usually they come from physiological experiments.  

The following figure shows an example of the data displayed in both geographical and environmental space.

<!--- Chunk for reading and ploting an example of each kind of data described above. Dependencies: mtime --->
```{r data, echo=F, cache=T, cache.extra=list(mtime("Background.csv"),mtime("SpD_50_20.csv"),mtime("tolerance_ranges_xy.csv"))}
back <- read.csv("Background.csv",header=T)
occu <- read.csv("SpD_50_20.csv",header=T)
tolran <- read.csv("tolerance_ranges_xy.csv",header=T)
plotdata(back=back,occ.sp=occu,tolran=tolran,sp.col="violetred3")
```

  The details about the assumptions and the statistical model are not provided here, however, it is important to notice that under this Bayesian framework the purpose is to simulate values from the posterior distribution of the parameters that define fundamental niche of a species. The resulting expression for the posterior comes from: (1) assuming an ellipsoidal shape for the niche (which translates into using a multivariate normal distribution to define the likelihood), and (2), the addition of a priori information for the centroid and variances describing the niche. The general expression of the posterior function is  
$$ f\left(\mu,A|D,E(t;G)\right) \propto \mathcal{L}\left(\mu,A\mid D\right)g_{1}\left(\mu\right)g_{2}\left(A\right),
$$  
where $\mathcal{L}\left(\mu,A\mid D\right)$ is the likelihood function and $g_{i}\left(\cdot\right)$ are the _a priori_ destities of the parameters of interest, $\mu$ and $A$. I already transformed this expression into R code and saved it as a function so that we can use it to calculate its values and to generate simulations from it:  
<!--- Chunk for displaying the code of the function Energy --->
```{r Energy_source, echo=F}
source("Energy.R")
Energy
```

  When working under a Bayesian framework, the posterior distribution is the objective function of the analysis since it comprises all the information about the parameters of interest.
  I use an MCMC algorithm called "t-walk" (which will be described in detail in the final project) to generate simulations from the posterior. The set of draws provides posterior values of the mean, $\mu$, and the precision matrix, $A$, which can be used to construct ellipsoids and to derive metrics reflecting the uncertainty in the estimated parameters.
  There is an R package that contains functions for running the t-walk algorithm, it is called "Rtwalk". I will use the function "Runtawlk" from this package to get simulations from the posterior. In order to do that, I need to define three main functions: Supp which checks that a new value in the Markov chain is valid, Initth which produces values from the _a priori_ distributions, and Energy which contains the posterior. I already defined and tested these functions as part of the assignement and, to illustrate, here is how I will use them:  
  <!--- Chunk for showing how the functions Supp, Initth and Energy will be used --->
```{r runMCMC, eval=F}
library(Rtwalk)
info<-Runtwalk(Tr=10000,dim=5,Obj=Energy,Supp=Supp,x0=Initth(),xp0=Initth())
}
```

  After simulating from the posterior, I will need another set of functions to analyze and display the results. As an example, I defined the function "PlotIterations.R" which takes the output of the function "Runtwalk" and plots the resulting ellipses in the environmental space.  
  In summary, as part of my research project, I am developing a model for niche estimation. Using this model to estimate niches involves a series of complex calculations that can be made in R. The visualization of the data and the results form the estimation requires as well of special functions. Up to know, I have the code for most of the functions needed to apply the model. As part of the final project of the class, I will work on making this functions as modular and reproducible as possible because as my research project advances, I will need to addapt them and reuse them. Once I have the final version of the model, it will then be easy to adapt the R package that I will develop for this class and get a version that I could share with people interested on estimating niches.