---
title: "Tests for the functions used in the final project"
author: "Laura Jiménez"
date: "March 2018"
output: html_document
fontsize: 11pt
---
## General notes
All the functions will be tested assuming that we are working with an environmental space of two dimensions. These functions can be generalized to work with higher-dimensional spaces (except for the ones used to display visualizations which only will work for two dimensions).
All the functions consider that mu (vector of means, or ellipses centroid) and A (precisión matrix, or inverse of the covariance matrix) are the parameters that define the multivariate normal distribution associated with the fundamental niche of the species. These two variables are global variables in the main program, so we need to be careful when using them in the functions.

## Tests for Supp

```{r Supp_source, echo=T}
source("Supp.R")
```
In order to test the function Supp we need to define the following variables:  
- Et = a matrix with two columns such that each row contains environmental combinations representing the climatic background (every time the values of mu and A change we need to check that they won't produce an error in future calculations when combined with the points of the background, see the expression of the posterior)  
- mu.lim = a vector that contains the minimum and maximum values for each of the entries of a vector representing the mean of a MVN distribution  
- th = a vector that contains the values of the mean of a MVN distribution and the unique elements from its precision matrix

We need to test different cases:  
1. mu belongs to the given ranges and A is a positive matrix = passed  
```{r Supp_test1, echo=T, cache=T}
# read the file that contains the climatic background points
env <- read.csv("VarsCrdsBackground.csv",header=T) ## provide a url instead!!
# only keep the columng that correspond to climatic data
Et <- as.matrix(env[ , c(6,7)])
# notice that every pair of element in mu.lim correspond to the range of each variable in the multivariate space
mu.lim <- c(0,1,-1,1)
mu <- c(0.69,-0.25)
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
# notice the order in which the entries of th are specified
th <- c(mu,A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```
2. mu belongs to the given ranges and A isn't a positive matrix = failed  
```{r Supp_test2, echo=T}
mu.lim <- c(0,1,-1,1)
mu <- c(0.69,-0.25)
# it is enough to change the sign of one entry in the previous matrix to get a matrix that is not possitive definite
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
th <- c(mu,-A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```
3. mu don't belong to the given ranges = failed  
```{r Supp_test3, echo=T}
mu.lim <- c(0,1,-1,1)
# it is enough to change one of the entries of mu to make Supp fail
mu <- c(0.69,-0.25) + 1
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
th <- c(mu,A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```

## Tests for Initth
The function Initth do not accept any arguments, however, there are three variables that must be defined before using this function: mu0, CholSigma0 and CholW. These are the parameters of the a priori distribution for mu and A.
```{r Initth_source, echo=T}
source("Initth.R")
```
After defining mu0 and A0, call the function and verify that the output is a numeric vector of length equals to 5.
```{r Inith_test1, echo=T,eval=F}
mu0 <- c(0.69,-0.25)
A0 <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
result <- Initth()
if(length(result)==5){
  print("passed")
  print(result)
} else {
  print("failed")
  print(result)
}
```

## Tests for PlotIterations
A tests will be defined once the function code is ready or we can provide a figure with an example of what we expect as an outpur so we can visually check if the function produced good results.
```{r PlotIterations_source, echo=T}
source("PlotIterations.R")
```

```{r PlotIterations_test1, echo=F}
print("test under construction")
```

## Tests for Energy
This function can be different if we are working with different models when we defined the likelihood function or the a priori distributions for the parameters of interest.
```{r Energy_source, echo=T}
source("Energy.R")
```
Instead of applying a test on the function, we can provide the mathematical expression of the objective function so we can verify that the code matches this expression.
```{r Energy_test1, echo=F}
print("test under construction")
```
