---
title: "Tests for the functions used in the final project"
author: "Laura Jiménez"
date: "March 2018"
output: html_document
fontsize: 11pt
---
## General notes

All the functions will be tested assuming that we are working with an environmental space of two dimensions. These functions can be modified easylt to work with higher-dimensional spaces (except for the ones used to display visualizations which only will work for two dimensions).
All the functions consider that mu (vector of means, or ellipses centroid) and A (precisión matrix, or inverse of the covariance matrix) are the parameters that define the multivariate normal distribution associated with the fundamental niche of the species. These two variables are global variables in the main program, so we need to be careful when using them in the functions.

## Tests for Supp

```{r Supp_source, echo=T}
source("Supp.R")
```

In order to test the function Supp we need to define the following variables:  
- Et = a matrix with two columns such that each row contains environmental combinations representing the climatic background (every time the values of mu and A change we need to check that they won't produce an error in future calculations when combined with the points of the background, see the expression of the posterior)  
- mu.lim = a vector that contains the minimum and maximum values for each of the entries of a vector representing the mean of a MVN distribution  
- th = a vector that contains the values of the mean of a MVN distribution and the unique elements from its precision matrix

We need to test three different cases:  
1. mu belongs to the given ranges and A is a positive matrix, so that the output must be TRUE  
```{r Supp_test1, echo=T, cache=T}
# read the file that contains the climatic background points
env <- read.csv("Background.csv",header=T) ## provide a url instead!!
# only keep the columng that correspond to climatic data
Et <- as.matrix(env[ , c(6,7)])
# notice that every pair of element in mu.lim correspond to the range of each variable in the multivariate space
mu.lim <- c(0,1,-1,1)
mu <- c(0.69,-0.25)
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
# notice the order in which the entries of th are specified
th <- c(mu,A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```
2. mu belongs to the given ranges and A isn't a positive matrix, so that the output must be FALSE  
```{r Supp_test2, echo=T}
mu.lim <- c(0,1,-1,1)
mu <- c(0.69,-0.25)
# it is enough to change the sign of one entry in the previous matrix to get a matrix that is not possitive definite
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
th <- c(mu,-A[1,1],A[2,2],A[1,2])

if(!Supp(th)){
  print("passed")
} else {
  print("failed")
}
```
3. mu don't belong to the given ranges, so that the output must be FALSE  
```{r Supp_test3, echo=T}
mu.lim <- c(0,1,-1,1)
# it is enough to change one of the entries of mu to make Supp fail
mu <- c(0.69,-0.25) + 1
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
th <- c(mu,A[1,1],A[2,2],A[1,2])

if(!Supp(th)){
  print("passed")
} else {
  print("failed")
}
```

## Tests for Initth

The function Initth do not accept any arguments, however, there are 
three variables that must be defined before using this function: mu0, CholSigma0 and CholW. These are the parameters of the a priori distribution for mu and A.

```{r Initth_source, echo=T}
source("Initth.R")
```
After defining mu0 and Sigma0, define the variables CholSigma0 and Cholw (the Cholesky decompositions of Sigma0 and W, respectively), then, call the function and verify that the output is a numeric vector of length equals to 5.
```{r Inith_test1, echo=T}
# Define the required variables
mu0 <- c(-1.09, -0.20) # mean vector, MVN prior
Sigma0 <- matrix(c(0.0144,0,0,0.0324),ncol=2,byrow=T) # covariance matrix, MVN prior
CholSigma0 <- chol(Sigma0)
alpha <- 2 #shape parameter, Wishart prior
W <- alpha*chol2inv(CholSigma0)  #scale matrix, Wishart prior
CholW <- chol(W)
# mu0, Sigma0 and CholW will be used when evaluating Initth
result1 <- Initth()
result2 <- Initth()
# use the function twice to check that it produces two vectors of length equals to 5 and that these vectors are different
if(length(c(result1,result2)) == 10){
  if(sum(result1 - result2) != 0)
    print("passed")
  print(round(result1,2))
  print(round(result2,2))
} else {
  print("failed")
  print(round(result1,2))
  print(round(result2,2))
}
```

## Tests for PlotIterations
A test will be defined once the function code is ready or we can provide a figure with an example of what we expect as an output so we can visually check if the function produced good results.
```{r PlotIterations_source, echo=T}
source("PlotIterations.R")
```

```{r PlotIterations_test1, echo=F}
print("test under construction")
```

## Tests for Energy

This function can be different if we are working with different models when we defined the likelihood function or the a priori distributions for the parameters of interest.  
```{r Energy_source, echo=T}
source("Energy.R")
Energy
```

Instead of applying a test on the function, I provide here the mathematical expression of the objective function so we can verify that the code matches this expression.
These two terms must be contained in the Energy function and correspond to the log-likelihood function of mu and A:  
$$ -\frac{1}{2}\sum_{i=1}^{n}\left(x_{i}-\mu\right)^{T}A\left(x_{i}-\mu\right) - n\log\left\{\sum_{y\in E(t;G)}\exp\left[-\frac{1}{2}\left(y-\mu\right)^{T}A\left(y-\mu\right)\right]\right\} $$  
Notice that the argument of the function $log$ is coded as _suma.Et_ inside the Energy function. The variable coded as _env.sp_ represents the matrix that contains the environmental combinations where the species was observed, $x_{1},...,x_{n}$.
The other three terms contained in the Energy function correspond to the _a priori_ densities:  
$$  -\frac{1}{2}(\mu-\mu_0)^{T}A_0(\mu-\mu_0) + \frac{\alpha-d-1}{2}\log(det(A)) - \frac{1}{2}tr(AW^{-1}), $$  
where $det(A)$ is the determinant of the matrix A and $tr(\cdot)$ is the trace of the matrix (the sum of the elements in the diagonal).
