---
title: "Tests"
author: "Laura Jiménez"
date: "March 2018"
output: html_document
fontsize: 11pt
geometry: margin=1in
documentclass: article
---

# Tests for Supp

```{r Supp_source, echo=T}
source("Supp.R")
```
In order to test the function Supp we need to define the following variables:
-mu.lim = a vector that contains the minimum and maximum values for each of the entries of a vector representing the mean of a MVN distribution
-mu =  the mean of a MVN distribution
-A = a square matrix of dimensions equal to the lenght of mu

We need to test different cases: 1) when mu is in the permited ranges and A is a positive matrix,...

```{r Supp_test1, echo=T}

```

# Tests for Initth

```{r Initth_source, echo=T}
source("Initth.R")
```

```{r Inith_test1, echo=T}

```

# Tests for PlotIterations

```{r PlotIterations_source, echo=T}
source("PlotIterations.R")
```

```{r PlotIterations_test1, echo=T}

```

# Tests for Energy

```{r Energy_source, echo=T}
source("Energy.R")
```

```{r Energy_test1, echo=T}

```
