---
title: "Tests"
author: "Laura Jiménez"
date: "March 2018"
output: html_document
fontsize: 11pt
geometry: margin=1in
documentclass: article
---
## General note
All the functions will be tested assuming that we are working with an environmental space of two dimensions. These functions can be generalized to work in a space of higher dimension,except for the ones used to display visualizations which only work for two dimensions.

## Tests for Supp

```{r Supp_source, echo=T}
source("Supp.R")
```
In order to test the function Supp we need to define the following variables:  
- mu.lim = a vector that contains the minimum and maximum values for each of the entries of a vector representing the mean of a MVN distribution  
- th = a vector that contains the values of the mean of a MVN distribution and the unique elements from its precision matrix

We need to test different cases:  
1. when mu belongs to the given ranges and A is a positive matrix (passed)  
```{r Supp_test1, echo=T,eval=F}
# notice that every pair of element in mu.lim correspond to the range of each variable in the multivariate space
mu.lim <- c(0,1,-1,1)
mu <- c(0.69,-0.25)
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
# notice that order in which the entries of th must be specified
th <- c(mu,A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```
2. when mu belongs to the given ranges and A isn't a positive matrix (failed)  
```{r Supp_test2, echo=T,eval=F}
mu.lim <- c(0,1,-1,1)
mu <- c(0.69,-0.25)
# it is enough to change the sign of one entry in the previous matrix to get a matrix that is not possitive definite
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
th <- c(mu,-A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```
3. when mu don't belong to the given ranges (failed)  
```{r Supp_test3, echo=T,eval=F}
mu.lim <- c(0,1,-1,1)
# it is enough to change one of the entries of mu to make Supp fail
mu <- c(0.69,-0.25) + 1
A <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
th <- c(mu,A[1,1],A[2,2],A[1,2])

if(Supp(th)){
  print("passed")
} else {
  print("failed")
}
```

## Tests for Initth
The function Initth do not accept any arguments, however, there are two variables that must be defined before using this function: mu0 and A0. These are the parameters of the a priori distribution for mu.
```{r Initth_source, echo=T}
source("Initth.R")
```
After defining mu0 and A0, call the function and verify that the output is a numeric vector of length equals to 5.
```{r Inith_test1, echo=T,eval=F}
mu0 <- c(0.69,-0.25)
A0 <- matrix(c(42.91354,-20.97932,-20.97932,23.58958),ncol=2,byrow=T)
result <- Initth()
if(length(result)==5){
  print("passed")
  print(result)
} else {
  print("failed")
  print(result)
}
```

## Tests for PlotIterations
A tests will be defined once the function code is ready or we can provide a figure with an example of what we expect as an outpur so we can visually check if the function produced good results.
```{r PlotIterations_source, echo=T}
source("PlotIterations.R")
```

```{r PlotIterations_test1, echo=F}
print("test under construction")
```

## Tests for Energy
This function can be different if we are working with different models when we defined the likelihood function or the a priori distributions for the parameters of interest.
```{r Energy_source, echo=T}
source("Energy.R")
```
Instead of applying a test on the function, we can provide the mathematical expression of the objective function so we can verify that the code matches this expression.
```{r Energy_test1, echo=F}
print("test under construction")
```
